---
title: "Final Project"
author: "Marisa Manno, Georgia Wright, Elizabeth Chin"
date: "12/1/2021"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, root.dir = "~/Desktop/Stat1261", 
                      out.width = "70%", fig.align = "center")
library(tidyverse)
library(dplyr)
library(RColorBrewer)
library(broom)
library(modelr)
```

```{r}

atus <- read.delim("atusact-0320/atusact_0320.dat", sep="," , header=T)
atus <- atus %>%
  format(atus$TUCASEID, scientific = FALSE) 

roster <- read.delim("atusrost-0320/atusrost_0320.dat", sep="," , header=T) 
roster<- roster %>%format(roster$TUCASEID, scientific = FALSE)


```

#### Variables:

We will be looking at how time spent on different categories of travel affects sleep,
so we have listed the 6 digit (TRCODEP) code for sleep along with the 4 digit (TRTIER2P) 
codes of each category of travel, each starting with 18 to indicate traveling for that reason. \ 

- 010101 sleeping \
- 18 travel: \
  - 1801 personal care
  - 1802 household activities 
  - 1803 travel related to caring for household members 
  - 1804 travel related to caring for nonhousehold members
  - 1805 work 
  - 1806 education
  - 1807 consumer purchases 
  - 1808 using professional and personal care services
  - 1809 household services
  - 1810 using government services and civic obligations 
  - 1811 eating and drinking
  - 1812 socializing relaxing leisure
  - 1813 sports exercise recreation
  - 1814 religious/spiritual
  - 1815 volunteer 
  - 1816 phone calls
  - 1818 security procedures related to traveling 
  - 1899 nec 



First, we filter our variables of interest from the atus data and select columns 
we need (i.e we did not include things like if a child was with you and where 
you were during the activity).
We have called this shortened data frame "atus_main"

```{r}
atus_main <- atus %>%
  filter(TRCODEP == 010101 | TRTIER1P == 18 | TRTIER2P == 0201 | TRTIER2P == 0501 |
          TRTIER2P == 0601 | TRTIER2P == 1203 ) %>%
  dplyr::select(TUCASEID, TRCODEP, TRTIER1P, TRTIER2P, TUACTDUR) 



```


#### Visualize each category of travel

atus_travel is our main data set with only the codes corresponding to travel.
First is a bar graph of each category of travel show in hours of time spent on 
that category.
```{r}
atus_travel = atus_main %>%
  filter(TRTIER1P == 18 )

head(atus_travel)

```
We are going to determine how many hours were spent for each travel related 
activity across households. We predict that the most time spent traveling 
will be related to work, education, and caring for household members. 
```{r}
travel_sum <- atus_travel %>% 
  group_by(TRTIER2P) %>% 
  summarise(min=sum(as.numeric(TUACTDUR)))  %>% 
  mutate(hours=min/60)


head(travel_sum)
```


```{r}
travel_sum$TRTIER2P <- as.factor(travel_sum$TRTIER2P)

colorCount = nrow(travel_sum)
getPalette = colorRampPalette(brewer.pal(9, "Set1"))

ggplot(travel_sum, aes(x=TRTIER2P, y=hours, fill=TRTIER2P)) +
geom_bar(stat = "identity")  + 
  theme(axis.text.x=element_text(angle = 90, hjust = 0)) + 
  ggtitle("Hours Spent Travelling by Category") +
  xlab("Travel Activity") + 
  ylab("Hours") +
  scale_fill_hue(name = NULL, labels = c("Personal Care", "Household Act", "Caring for Household Members", 
                         "Caring for NonHousehold Members", "Work", "Education", 
                         "Consumer Purchases", "Prof and Personal Care Service", 
                         "Household Services", "Gov Services and Civic Obligations",
                         "Eating and Drinking", "Socializing/Leisure", "Sports/Exercise",
                         "Religious/Spiritual", "Volunteer", "Phone Calls", 
                         "Security Procedures", "n.e.c.")) 

  
```



This bar chart is a visualization of total hours spent for each travel activity 
across all households. Overall, the most time spent traveling is for consumer 
purchases. The second most is time spent traveling is for work followed by 
socializing and leisure.


#### Use the ATUS Roster file to visualize travel activities based on sex and age (18 and older)
```{r}
roster$TEAGE <- as.numeric(roster$TEAGE)
roster1<- roster %>% filter(TEAGE>=18)
rost_act<- roster1 %>% left_join(atus_main, by="TUCASEID")

```

Create variables for age_group and sex:
```{r}
rost_act1<-rost_act %>% mutate(age_group = ifelse(TEAGE <=25, "18-25", 
                                 ifelse(TEAGE<=35, "26-35", 
                                 ifelse(TEAGE<=45, "36-45", 
                                 ifelse(TEAGE<=55, "46-55",
                                ifelse(TEAGE<=65, "56-65",
                                ifelse(TEAGE<=75, "66-75",
                                ifelse(TEAGE<=85, "76-85",))))))))
head(rost_act1)

rost_act2<- rost_act1 %>% mutate(sex= ifelse(TESEX=="2", "Female", "Male"))
head(rost_act2)


```

Filter to display only travel activities and select Case ID, Activity, Activity \ 
duration, Age Group, and Sex.
```{r}
travel2<- rost_act2 %>% 
  filter(TRTIER1P == 18) %>% 
  group_by(TUCASEID) %>% 
  dplyr::select(TUCASEID, TRTIER1P, TRTIER2P, TUACTDUR, age_group, sex)
head(travel2) 
```
Calculate hours spent traveling for each age group. We hypothesize that 36-45 
year olds will spend the most time traveling because of work and childcare 
responsibilities.
```{r, warning = FALSE }
travel_age<-travel2 %>% 
  group_by(age_group) %>% 
  summarise(sum=sum(as.numeric(TUACTDUR))) %>% 
  mutate(hours=sum/60)

travel_age

ggplot(travel_age, aes(x=age_group, y=hours, fill=age_group))+
  geom_bar(stat = "identity")  + 
  theme(axis.text.x=element_text(angle = 90, hjust = 0)) + 
  ggtitle("Hours Spent Travelling For Age Groups") +
  xlab("Age Group") + ylab("Hours") +
  scale_fill_discrete(name = "Age Group") +
  scale_fill_brewer(palette = "Dark2") +
  guides(fill=guide_legend(title= "Age Group"))
  
```




The age-group 36-45 spends the most time traveling.

##### Travel activities of 36-45 year olds \ 
```{r}
group3<-travel2 %>% 
  filter(age_group=="36-45") %>% 
  group_by(TRTIER2P) %>% 
  summarise(sum=sum(as.numeric(TUACTDUR))) %>% 
  mutate(hours=sum/60) 

head(group3)
group3$TRTIER2P <- as.factor(group3$TRTIER2P)

ggplot(group3, aes(x=TRTIER2P, y=hours, fill=TRTIER2P)) +
  geom_bar(stat = "identity")  + 
  theme(axis.text.x=element_text(angle = 90, hjust = 0)) + 
  ggtitle("Hours Spent Travelling For 35-46 Year Olds") +
  xlab("Travel Activity") + ylab("Hours") +
  scale_fill_hue(name = NULL, labels = c("Personal Care", "Household Act", "Caring for Household Members", 
                         "Caring for NonHousehold Members", "Work", "Education", 
                         "Consumer Purchases", "Prof and Personal Care Service", 
                         "Household Services", "Gov Services and Civic Obligations",
                         "Eating and Drinking", "Socializing/Leisure", "Sports/Exercise",
                         "Religious/Spiritual", "Volunteer", "Phone Calls", 
                         "Security Procedures", "n.e.c.")) 
```




Individuals aged 35-46 spend the most time traveling for work and consumer 
purchases, which is consistent with the overall population.

##### Visualize time spent traveling for Females and Males \
We are looking to determine if differences in time traveling activities exist
based on sex.
```{r}
travel3<- travel2 %>% 
  group_by(TRTIER2P, sex) %>% 
  summarise(sum=sum(as.numeric(TUACTDUR))) %>% 
  mutate(hours=sum/60)

travel3

travel4<- travel2 %>% 
  group_by(sex) %>% 
  summarise(sum=sum(as.numeric(TUACTDUR))) %>% 
  mutate(hours=sum/60)
travel4

travel3$TRTIER2P <- as.factor(travel3$TRTIER2P)

ggplot(travel3, aes(x=TRTIER2P, y=hours, fill=TRTIER2P)) +
  geom_bar(stat = "identity")  + 
  theme(axis.text.x=element_text(angle = 90, hjust = 0)) + 
  ggtitle("Hours Spent Travelling Based on Sex") +
  xlab("Travel Activity") + ylab("Hours") + 
  facet_wrap("sex") +
  scale_fill_hue(name = NULL, labels = c("Personal Care", "Household Act", "Caring for Household Members", 
                         "Caring for NonHousehold Members", "Work", "Education", 
                         "Consumer Purchases", "Prof and Personal Care Service", 
                         "Household Services", "Gov Services and Civic Obligations",
                         "Eating and Drinking", "Socializing/Leisure", "Sports/Exercise",
                         "Religious/Spiritual", "Volunteer", "Phone Calls", 
                         "Security Procedures", "n.e.c.")) +
  theme(axis.ticks = element_blank(), 
        axis.text.x = element_blank()) 
  
```




Overall, females spend more time traveling than males. Females spend more time 
than males for every travel activity besides travel related to sports, 
exercise, and recreation. 


#### Scatter plot and linear model of Travel Categories on Sleep \ 
First, we made data frames of each of the three travel categories we are 
interested in, along with one for sleep.

1807 Consumer Purchases
```{r}
atus_1807 <- atus_travel %>%
  group_by(TUCASEID) %>%
  filter(TRTIER2P == 1807) %>%
  summarise(sum_1807=sum(as.numeric(TUACTDUR))) 

head(atus_1807)


```

1805 Work
```{r}
atus_1805 <- atus_travel %>%
  group_by(TUCASEID) %>%
  filter(TRTIER2P == 1805) %>%
  summarise(sum_1805 =sum(as.numeric(TUACTDUR)))

head(atus_1805)

```

1812 Socializing, Relaxing, Leisure
```{r}

atus_1812 <- atus_travel %>%
  group_by(TUCASEID) %>%
  filter(TRTIER2P == 1812) %>%
  summarise(sum_1812=sum(as.numeric(TUACTDUR)))

head(atus_1812)

```

Sleep
```{r}
atus_sleep = atus_main %>%
  filter(TRCODEP == 010101) %>% 
  group_by(TUCASEID) %>% 
  dplyr::select(TUCASEID, TUACTDUR) %>% 
  summarise(sum_sleep=sum(as.numeric(TUACTDUR))) 

head(atus_sleep)
```


Plot of Consumer Purchases on Sleep (1807)
```{r, warning = FALSE}
reg_1807 <- left_join(atus_1807, atus_sleep, by = "TUCASEID")
head(reg_1807)


reg_1807 %>% ggplot(aes(x= sum_1807, y= sum_sleep)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm")


```




There is a slight negative relationship between the variables, so we will
make a linear model. 
```{r}
travel_1807_mod <- lm(sum_sleep~sum_1807, data = reg_1807)
summary(travel_1807_mod)


confint(travel_1807_mod, "sum_1807", data = reg_1807, level = 0.99 )
```
We are 99% confident that the coefficient for 1807 (consumer purchases) is 
between -0.0985558 and -0.03186303. The 99% confidence interval does not 
contain 0, so we are 99% sure that there is a small negative effect. \ 

 
Let's look at our next variable
traveling for work (1805). 
```{r, warning = FALSE}

reg_1805 <- left_join(atus_1805, atus_sleep, by = "TUCASEID")
head(reg_1805)


reg_1805 %>% ggplot(aes(x= sum_1805, y= sum_sleep)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm") 
```

It looks like there is a negative correlation, but it seems to be caused by
high outliers on the x-axis, so we will show a scatter plot that eliminates
these points from view. 

```{r, warning = FALSE}
reg_1805 %>% ggplot(aes(x= sum_1805, y= sum_sleep)) +
  geom_point(alpha = 0.3) +
  scale_x_continuous(limits = c(0, 1000)) +
  geom_smooth(method = "lm")

```




Without the very high outliers we can still see a negative correlation. 

Regression of Time Traveling for work (1805) on Time Sleeping
```{r}
travel_1805_mod <- lm(sum_sleep~sum_1805, data = reg_1805)
summary(travel_1805_mod)

confint(travel_1805_mod, "sum_1805", data = reg_1805, level = 0.99 )


```
We are 99% confident that the coefficient for 1805 (work) is between -0.4071852 
and -0.3318718. Therefore there is a negative relationship because the 
confidence interval does not contain the value 0. 

Now we will look at travel time for relaxing and leisure (1812)
```{r, warning = FALSE}
reg_1812 <- left_join(atus_1812, atus_sleep, by = "TUCASEID")

reg_1812 %>% ggplot(aes(x= sum_1812, y= sum_sleep)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm")

```





There appears to be only a very slight relationship along with 3 high outliers.
The data without the outliers shows no relationship.
```{r}
reg_1812 %>% ggplot(aes(x= sum_1812, y= sum_sleep)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm") +
  scale_x_continuous( limits = c(0, 1100))

```





Now we will run a model with both 1805 (work) and 1807 (consumer purchases)
```{r}

travel_full_reg<- reg_1805 %>% left_join(reg_1807)
na.omit(travel_full_reg)

travel_full_mod <- lm(sum_sleep ~ sum_1805 + sum_1807, data=travel_full_reg)
summary(travel_full_mod)

confint(travel_full_mod, "sum_1805", data = travel_full_reg, level = 0.99 )
confint(travel_full_mod, "sum_1807", data = travel_full_reg, level = 0.99 )

```
Both coefficients have very low p-values and are significant. We are 99% confident 
that the 1805 (work) coefficient is between -0.5118716 and -0.3629791 and that 
the 1807 (consumer purchases) coefficient is between -0.3484959 and -0.172871. 
Therefore, we are 99% confident that both variables have a negative effect on 
sleep time. 


#### Part One Data Sets 
1. Marisa Manno - Both Republicans And Democrats Have an Age Problem (congress-terms.csv)
2. Georgia Wright - How Boomers Get High (drug-use-by-age.csv)
3. Elizabeth Chin - Dear Mona, Which State Has The Worst Drivers? 
    (bad_drivers from the fivethirtyeight library)

#### Group Contributions 
- Marisa Manno
  - created travel_sum data set 
  - created data frames for each of the top three travel categories and sleep
  - created simple linear regressions for the two correlated travel variables on sleep
  - created a model of both variables on sleep 
  - ran confidence intervals for each model
- Georgia Wright
  - bar graph for Hours Spent Traveling by Category
  - Wrangled the roster file 
  - created variables for age and sex
  - created other travel data sets for other graphs 
  - created graph for age group
  - created graph for 35-46 year olds
  - created graph for 35-46 year olds by sex
- Elizabeth Chin
  - created atus_main data set 
  - created atus_travel data set
  - added labels to graphs and themes
  - added labels to graph legends 
  - created scatter plots for the top three travel categories 

